Newly Left-Behind Children's Educational Performance Drops Slightly After Parental Migration

### 2023-11-09 MScThesisText.tex
Add latest literature

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{labor migration, left-behind children, educational performance\newline\indent Word count: 6,789}
\usepackage{csquotes}
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\typesectitle}}

@@

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Newly Left-Behind Children's Educational Performance Drops Slightly After Parental Migration},
  pdfauthor={Jiancheng Gu1},
  pdflang={en-EN},
  pdfkeywords={labor migration, left-behind children, educational performance},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

This paper's open science repository: \url{https://github.com/jianchenggu/MScThesis}
An earlier version of this article was submitted in partial fulfillment of the requirements for the degree -- Master of Science in Societal Resilience (Research).

@@

\affiliation{\vspace{0.5cm}\textsuperscript{1} Faculty of Social Sciences, Vrije Universiteit Amsterdam}

@@

May 15, 2023

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

How are stay-behind children faring in the absence of parents? Research findings are mixed due to the differential effects of parental migration. Consider children's education. According to the review articles by Venta and Cuervo (2022) and Van Hook and Glick (2020), labor migrants send remittances back, which can fund children's schooling as well as improve their nutritional and living conditions. To what degree can the inflow of remittances compensate for the loss of parental attention? Parental migration decreases the amount and quality of parental guidance, protection, and support. It is uncertain how these two opposing effects together shape children's educational performance. This uncertainty merits empirical investigation.

A challenge arises in estimating the effect of parent's migration on children's educational achievement. Experiments can unveil causal relations. It is unfeasible and unethical to design an experiment that randomly assigns parents to migrate without their children. Alternatively, a quasi-experimental design can help with causal inference. It requires that the cause be on par with random assignment to participants, conditional on identification assumptions.

@@

\hypertarget{chinas-context}{%
\subsection{China's Context}\label{chinas-context}}

Three features make China a suitable case for studying left-behind children. The most prominent feature is the huge scale of internal labor migration. According to the National Bureau of Statistics, the year 2021 recorded 292.5 million in-country migrant workers. They form the backbone of essential industries, such as retail, transportation, and manufacturing. The 2015 one-percent National Population Sample Survey estimated that 68.7 million Chinese children were left behind (NBS, 2022; NBS et al., 2017).

@@

The institutional discrimination against labor migrants finds its roots in China's ``multilevel citizenship'' system. As Vortherms (2021) explained, the constituents of citizenship -- ``membership rules and rights entitlements'' -- operate below the national level in China. The local state confers most of the rights and defines who is eligible to these rights through \emph{hukou}, or the household registration system:

\begin{quote}
A political institution rooted in centuries of Chinese bureaucracy, the modern hukou is the primary identity document defining membership and entitlements for local citizenship. Accessing fundamental citizenship rights depends on your registration status \ldots{} Registration includes basic identifying information, a specific address of registration, and a type: urban, rural, or resident. (Vortherms, 2021)
\end{quote}

In other words, the hukou system binds one's citizenship rights with one's place of registration. In theory, Chinese citizens can migrate freely within the country. In practice, they cannot change hukou freely. When rural residents move to the city, it takes months -- if not years -- to apply for a new hukou. Before receiving urban hukou, they can hardly send their children into public schools in the host city (Wan \& Vickers, 2022; Zhang, 2018). Migrant workers have organized petitions, strikes, and protests. These actions often fell on deaf ears and sometimes even ended with police crackdown (Chan, 2019; Göbel, 2021). Facing financial and policy constraints, many migrant laborers have to leave children in their hometowns. Only during the winter and summer breaks can these children reunite with parents.

How does China counter the issue of labor migration and left-behind children? According to X. Wang (2020) and Dong and Goodburn (2020), the government has gradually liberalized the multilevel citizenship scheme. It has introduced residence permit for internal migrants as the basis for social entitlements. First-tier cities like Beijing and Shanghai have brought forth a point-based system to grant hukou to newcomers. The system's evaluation criteria include in-demand talents and skills, educational and professional qualifications, and consecutive years of employment and tax-payment. This parallels how European states differentially incorporate foreign immigrants into national welfare regimes (Pasquali, 2022). Another policy response, as stated by Roberts and Hannum (2018), is the centralized boarding school. The Ministry of Education promotes boarding schools in the rural county-seats. These schools serve as a cost-effective way to enroll students in surrounding villages and to look after the left-behind children. In reality, such schools ``are often ill-equipped to ensure the healthy upbringing and psychological wellbeing of children'' (p.~6).

Social scientists have researched extensively about the education of Chinese left-behind children, but the current literature still has three major shortcomings. The first shortcoming lies in sampling. For many studies, the geographic coverage is limited to one to a few counties; The participants are children left behind in rural areas but not those in urban areas (e.g. Hu, 2019; Zhou et al., 2014). Alpermann (2020) pointed that the Chinese government reinforces ``an urban hierarchy'' besides the rural-urban divide. In the hierarchy's bottom are small- and medium-sized cities in the inland. They also see parents migrating to metropolises and leaving children behind (S. Wang, 2023). The National Bureau of Statistics noted that the number of urban left-behind children have been growing, and the trend is expected to continue (NBS et al., 2017). The second shortcoming lies in research design. Some studies rely on cross-sectional data that focus on single time points (e.g. Khalid et al., 2022). These studies fail to capture the dynamic nature of migratory journey. Some other studies probe correlation from panel data using basic ordinary least squares regression (e.g. L. Chen et al., 2023). These studies could have employed causal inference techniques. The third shortcoming lies in the measurement of outcome. Chang et al. (2019) and L. Wang et al. (2019) measured students' educational performance with the standardized test score of only one academic subject, namely mathematics. To tackle these limitations, I adopted a nationally representative two-wave panel dataset. My sample includes urban left-behind children as well. I included the test results of multiple subjects to measure comprehensive academic achievement.

\hypertarget{theory-and-hypotheses}{%
\subsection{Theory and Hypotheses}\label{theory-and-hypotheses}}

Between parental migration and children's education, the mechanisms trace to two theoretical frameworks. One is the resource generation model. It holds that parental migration brings forth greater resources in the form of economic capital (Brauw, 2019; Shen et al., 2021). Economic capital is, in Bourdieu's words, ``immediately and directly convertible into money and may be institutionalized in the form of property rights'' (2002, p. 16). Economic resources can help individuals acquire human capital, which is the stock of ``knowledge, skills, and health that people accumulate over their lives'' (World Bank, 2020, p. 1). Yang (2008) noted that in the Philippines, a sharp depreciation of peso increased the amount of remittances from migrants working abroad. This subsequently boosted children's school enrollment and educational spending. Vikram (2021) performed propensity score matching and fixed-effect regression on a two-wave panel dataset featuring Indian children aged 8-11. She showed that fathers' migration was linked with increases in left-behind children's reading capability and educational expenditure (p.~991). In China, Gu (2022) interviewed migrant workers and their family members. She found that they aspire to move up the social ladder; Such aspiration translates into ``intensive and all-out investment in children's education'' (p.~526). Oftentimes, labor migration presented the only way to sustain their children's education and livelihood. Peng (2018) interviewed first-generation migrant mothers whose initial migration dated back to the 1990s. To them, ``good mothering'' primarily meant offering economic support for their children in education, which can help their children climb the ladder of success.

The other theoretical model is the family disruption model (Shen et al., 2021). Children flourish under parental nurturing care. This term refers to the ``stable environments that promote health and adequate nutrition, protect from threats \ldots{} opportunities for learning,'' and ``responsive, emotionally supportive and developmentally enriching relationships'' (Black et al., 2021, p. 1). Parental migration decreases nurturing care, as characterized by less homework supervision and emotional support by parents. Hong talked to a group of left-behind children at a rural school (Hong \& Fuller, 2019). These students expressed the utter sense of loneliness ``with no advocate, no support and with no one looking out for them'' (p.~13). Many of them felt indifferent to grades and disengaged with school. Some even longed to quit school to pursue a free, independent life (p.~14). Murphy (2022) talked to left-behind adolescents at a Chinese village. Some of them ``embraced the `left-behind' label'' because it captured their need for more parental care (p.~185), ``especially companionship and affection in daily life'' (p.~186). According to C. Chen (2023), caregiver-child interactions are infrequent and unstable for left-behind children. She showed that the loss in such interactions ``mediates almost all of the adverse effects of parental absence.'' Her conclusion was based on structural equation modeling using a cross-sectional survey with a sample of Chinese children aged three to fifteen.

Given the two opposing effects, researchers have obtained mixed results on how parental migration impacts children's education. Botezat and Pfeiffer (2020) adopted past immigration rate as an instrumental variable. They found a positive effect of parents' migration on academic grades of adolescents aged eleven to fifteen in Romania. The team of L. Wang et al. (2019) collected data from junior high school students in central China. Their two-way fixed-effect model showed no significant effect of parental migration on left-behind children's achievement in mathematical test. Nguyen (2016) conducted fixed-effect regression and demonstrated that parental migration was associated with lower cognitive capabilities among Indian and Vietnamese children aged five to eight.

@@

Researchers have studied whether the impact differs between mother's migration and father's migration. Xu et al. (2019) adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had ``little or no association with negative outcomes'' on children's academic and cognitive performances. Only-mother absence, however, showed a ``strong association with negative outcomes'' (p.~1646). Chen et al. (2019) conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that maternal migration was ``negatively associated with children's social competence and academic performance,'' while paternal migration had no direct effects on the two outcomes (pp.~860-861).

Murphy (2020) noted that in some Asian and Latin American societies, the social norms designate men as breadwinners and women as caregivers. Under such norms, fathers' migration is more likely to win children's acceptance than does mothers' migration. Left-behind children may even resent their migrant mothers for transgressing on the expected care-giving roles (Murphy, 2020). I propose the second hypothesis:

@@

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{sampling}{%
\subsection{Sampling}\label{sampling}}

I obtained the data from the China Education Panel Survey (CEPS) (NSRC, n.d.). It is a nationally representative, school-based survey featuring junior high school students. Currently, two waves of data are available. The CEPS team operates under the National Survey Research Center at the Renmin University of China. The survey team adopted multi-stage stratified probability proportional to size sampling and administered a paper-and-pencil questionnaire to each participant. The wave one baseline survey took place in the 2013-2014 academic year with a sample size of about twenty thousand students in the seventh and ninth grades. They were nested in 438 classrooms of 112 schools in 28 urban districts or rural counties (hereafter `counties'). The wave two follow-up survey in the 2014-2015 academic year lost track of the ninth grade cohort and only surveyed 10279 of the seventh grade cohort. I further restricted the sample to students living with both parents at the original domicile (place of hukou registry) at the baseline survey. To tackle missing values, I adopted list-wise deletion for the dependent, independent, and control variables. The final sample size amounts to 4114.

\hypertarget{measurement}{%
\subsection{Measurement}\label{measurement}}

@@

The lists of covariates are in Table~\ref{tab:tab-ctvar-dit} -- Table~\ref{tab:tab-ctvar-msr} in the Results section. At the children's level, I controlled for age, sex, self-rated health, ethnicity, hukou type, number of siblings, whether the child attends boarding school, and whether the child had skipped grade or repeated grade. Age, sex, and health are associated with educational performance (Gu \& Yeung, 2021). Regarding ethnicity and hukou status, Han Chinese and urban hukou holders outperform ethnic minorities and rural hukou holders at school (Hasmath, 2022; Huang et al., 2020; D. Xu et al., 2018). I controlled for boarding at school because research results vary regarding the effect of boarding at school on academic performance (S. Guo et al., 2021; M. Liu \& Villa, 2020). I also included the number of siblings as a covariate, considering the quantity-quality trade-off in multi-child family (R. Guo et al., 2022). At the household level, I controlled for the levels of education of father and mother, whether the parent has a white collar job, whether the household has internet access, the perceived household financial conditions, and the number of extracurricular books at home. These variables measure family socioeconomic status, which is related to children's educational achievement (J. Liu et al., 2020; Young \& Hannum, 2018). Finally, I controlled for the level of education that the child aspires to achieve and the parental expectation on the child's level of education. Children's academic achievement, their educational aspiration, and their parents' expectations are interrelated (Chen et al., 2023; Xu \& Montgomery, 2021). All control variables were measured in the baseline survey.

\hypertarget{analytical-strategy}{%
\subsection{Analytical Strategy}\label{analytical-strategy}}

@@

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

@@

\hypertarget{appendix-the-appendix}{%
\appendix}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\makeatletter
\usepackage{etoolbox}
\patchcmd{\maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{labor migration, left-behind children, educational performance\newline\indent Word count: 6,000}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\typesectitle}}

@@

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Newly Left-Behind Children's Educational Performance Drops Slightly After Parental Migration},
  pdfauthor={Jiancheng Gu1},
  pdflang={en-EN},
  pdfkeywords={labor migration, left-behind children, educational performance},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

This paper's open science repository: \url{https://github.com/jchgu/MScThesis}

Correspondence concerning this article should be addressed to Jiancheng Gu, Dr.-Karl-Dorrek-Strasse 30, 3500 Krems, Austria. E-mail: \href{mailto:jiancheng.gu@donau-uni.ac.at}{\nolinkurl{jiancheng.gu@donau-uni.ac.at}}

@@

\affiliation{\vspace{0.5cm}\textsuperscript{1} Department for Migration and Globalisation, Universität für Weiterbildung Krems}

@@

November 09, 2023

@@

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

@@

How are stay-behind children faring in the absence of parents? Research findings are mixed due to the differential effects of parental migration. Consider children's education. According to Venta and Cuervo (2022) and Van Hook and Glick (2020), labor migrants send remittances back, which can fund children's schooling as well as improve their nutritional and living conditions. To what degree can the inflow of remittances compensate for the loss of parental attention? Parental migration decreases the amount and quality of parental guidance, protection, and support. It is uncertain how these two opposing effects together shape children's educational performance. This uncertainty merits empirical investigation.

Quantitative social scientists have researched extensively about the education of left-behind children, but recent literature still has shortcomings. Some studies fail to capture the dynamic nature of migratory process as they draw upon cross-sectional data that focus on single time points (e.g. Ahmed et al., 2023; Gong \& Rao, 2023; Viet Nguyen \& Hoang Vu, 2023). Some other studies only probe correlation using regression analysis (e.g. L. Chen et al., 2023; Huang et al., 2020; R. Liu \& Hannum, 2023). These studies could have employed causal inference techniques.

@@

\hypertarget{chinas-context}{%
\subsubsection{China's Context}\label{chinas-context}}

Three features make China a suitable case for studying left-behind children. The most prominent feature is the huge scale of internal labor migration. According to the National Bureau of Statistics, the year 2021 recorded 292.5 million in-country migrant workers. They form the backbone of essential industries, such as retail, transportation, and manufacturing. The 2015 one-percent National Population Sample Survey estimated that 68.7 million Chinese children were left behind (NBS et al., 2017; NBS, 2022).

@@

The institutional discrimination against labor migrants finds its roots in China's ``multilevel citizenship'' system. As Vortherms (2023) explained, the constituents of citizenship -- ``membership rules'' and ``rights entitlements'' -- operate below the national level in China. The local state confers most of the rights and defines who is eligible to these rights through \emph{hukou}, or the household registration system. The system differentiates people into local versus external and urban versus rural. In the rural-urban divide, urban hukou holders enjoy greater access to higher-quality welfare benefits and public services compared to rural hukou holders. In the local-external divide, local hukou holders are those registered in the same county where they live. In theory, Chinese citizens can migrate freely within the country. In practice, they cannot change hukou freely. When rural residents move to the city, it takes months -- if not years -- to apply for a new hukou. Before receiving urban hukou, they can hardly send their children into public schools in the host city (Y. Chen \& Fu, 2023; Zhang, 2018). These institutions turn rural migrants into second-class citizens in effect. Migrant workers have organized petitions, strikes, and protests. These actions often fell on deaf ears and sometimes even ended with police crackdown (Chan, 2019; Göbel, 2021). Facing financial and policy constraints, many migrant laborers have to leave children in their hometowns. Only during the winter and summer breaks can these children reunite with parents.

How does China counter the issue of labor migration and left-behind children? According to Dong and Goodburn (2020), the government has gradually liberalized the multilevel citizenship scheme. It has introduced residence permit for internal migrants as the basis for social entitlements. First-tier cities like Beijing and Shanghai have brought forth a point-based system to grant hukou to newcomers. The system's evaluation criteria include in-demand talents and skills, educational and professional qualifications, and consecutive years of employment and tax-payment. This parallels how European states differentially incorporate foreign immigrants into national welfare regimes (Pasquali, 2022). Another policy response, as stated by Roberts and Hannum (2018), is the centralized boarding school. The Ministry of Education promotes boarding schools in the rural county-seats. These schools serve as a cost-effective way to enroll students in surrounding villages and to look after the left-behind children. In reality, such schools ``are often ill-equipped to ensure the healthy upbringing and psychological wellbeing of children'' (p.~6).

\hypertarget{theory-and-hypotheses}{%
\subsubsection{Theory and Hypotheses}\label{theory-and-hypotheses}}

Between parental migration and children's education, the mechanisms trace to two theoretical frameworks. One is the resource generation model. It holds that parental migration brings forth greater resources that can help children acquire human capital -- the stock of ``knowledge, skills, and health that people accumulate over their lives'' (World Bank, 2020, p. 1). Yang (2008) noted that in the Philippines, a sharp depreciation of peso increased the amount of remittances from migrants working abroad. This subsequently boosted children's school enrollment and educational spending. Vikram (2021) performed propensity score matching and fixed-effect regression on a two-wave panel dataset featuring Indian children aged eight to eleven. She showed that fathers' migration was linked with increases in left-behind children's reading capability and educational expenditure (p.~991). In China, Gu (2022) interviewed migrant workers and their family members. She found that they aspire to move up the social ladder; Such aspiration translates into ``intensive and all-out investment in children's education'' (p.~526). Oftentimes, labor migration presented the only way to sustain their children's education and livelihood. Peng (2018) interviewed first-generation migrant mothers whose initial migration dated back to the 1990s. To them, ``good mothering'' primarily meant offering economic support for their children in education, which can help their children climb the ladder of success.

The other theoretical model is the family disruption model (Shen et al., 2021). Children flourish under parental nurturing care. This term refers to the ``stable environments that promote health and adequate nutrition, protect from threats \ldots{} opportunities for learning,'' and ``responsive, emotionally supportive and developmentally enriching relationships'' (Black et al., 2021, p. 1). According to C. Chen (2023), caregiver-child interactions are infrequent and unstable for left-behind children. The loss in such interactions ``mediates almost all of the adverse effects of parental absence.'' This conclusion was based on structural equation modeling using a cross-sectional survey with a sample of Chinese children aged three to fifteen. Murphy (2022) talked to children left behind in a Chinese village. Some of them ``embraced the `left-behind' label'' because it captured their need for more parental care (p.~185), ``especially companionship and affection in daily life'' (p.~186). In Romania, Udrea and Guiu (2022) and Diaconu-Gherasim et al. (2023) interviewed left-behind adolescents. Many of them felt disorientated about the lack of parents' support and oversight in schoolwork.

Given the two opposing effects, researchers have obtained mixed results on how parental migration impacts children's education. Botezat and Pfeiffer (2020) adopted past immigration rate as an instrumental variable. They found a positive effect of parents' migration on academic grades of adolescents aged eleven to fifteen in Romania. L. Wang et al. (2019) collected data from junior high school students in central China. Their two-way fixed-effect model showed no significant effect of parental migration on left-behind children's achievement in mathematical test. Chae and Glick (2019) analyzed the Migration and Remittances Household Survey of Kenya, Senegal, and Burkina Faso. They observed significant correlation between parental migration and left-behind children's school enrollment in Kenya, but not in the other two West African countries. Nguyen (2016) conducted fixed-effect regression and demonstrated that parental migration was associated with lower cognitive capabilities among Indian and Vietnamese children aged five to eight.

@@

Researchers have studied whether the impact differs between mother's migration and father's migration. R. Liu and Hannum (2023) analyzed how fifteen-year-old students' academic performance is associated with parental absence across 59 countries, using the 2012 Program of International Student Assessment data. Compared to students living with both parents, students perform worse in academic tests when only mother is absent or both parents are absent; When only father is absent, however, students are not significantly below par. Xu et al. (2019) adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had ``little or no association with negative outcomes'' on children's academic and cognitive performances. Only-mother absence, however, showed a ``strong association with negative outcomes'' (p.~1646). Chen et al. (2019) conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that maternal migration was ``negatively associated with children's social competence and academic performance,'' while paternal migration had no direct effects on the two outcomes (pp.~860-861).

Murphy (2020) and Fan (2023) noted that in some Asian and Latin American societies, the social norms designate men as breadwinners and women as caregivers. Under such norms, fathers' migration is more likely to win children's acceptance than does mothers' migration. Left-behind children may even resent their migrant mothers for transgressing on the expected care-giving roles. I propose the second hypothesis:

@@

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

\hypertarget{sampling}{%
\subsubsection{Sampling}\label{sampling}}

I obtained the data from the China Education Panel Survey (CEPS) (NSRC, n.d.). It is a nationally representative, school-based survey featuring junior high school students. Currently, two waves of data are available. The CEPS team operates under the National Survey Research Center at the Renmin University of China. The survey team adopted multi-stage stratified probability proportional to size sampling and administered a paper-and-pencil questionnaire to each participant. The wave one baseline survey took place in the 2013-2014 academic year with a sample size of about twenty thousand students in the seventh and ninth grades. They were nested in 438 classrooms of 112 schools in 28 urban districts or rural counties (hereafter `counties'). The wave two follow-up survey in the 2014-2015 academic year lost track of the ninth grade cohort and only surveyed 10279 of the seventh grade cohort. I further restricted the sample to students living with both parents at the original domicile (place of hukou registry) at the baseline survey. To tackle missing values, I adopted list-wise deletion for the dependent, independent, and control variables. The final sample size amounts to 4204.

\hypertarget{measurement}{%
\subsubsection{Measurement}\label{measurement}}

@@

The lists of covariates are in Table 1 -- Table 5 in the Results section. At the children's level, I controlled for age, sex, self-rated health, ethnicity, hukou type, number of siblings, whether the child had attended preschool, and whether the child attends boarding school. Age, sex, and health are associated with educational performance (Gu \& Yeung, 2021). Regarding ethnicity and hukou status, Han Chinese and urban hukou holders outperform ethnic minorities and rural hukou holders at school (Hasmath, 2023; Huang et al., 2020; D. Xu et al., 2018). Attending preschool in early childhood and enrolling in boarding school in adolescence are predictors of educational performance (Guo et al., 2021; M. Liu \& Villa, 2020; Zheng et al., 2021). I also included the number of siblings as a covariate, considering the quantity-quality trade-off in multi-child family (Doepke, 2015). At the household level, I controlled for the levels of education of father and mother, whether the parent has a white collar job, whether the household has internet access, the perceived household financial conditions, and the number of extracurricular books at home. These variables measure family socioeconomic status, which is related to children's educational achievement (J. Liu et al., 2020; Young \& Hannum, 2018). Finally, I controlled for the level of education that the child aspires to achieve and the parental expectation on the child's level of education. Children's academic achievement, their educational aspiration, and their parents' expectations are interrelated (Chen et al., 2023; Xu \& Montgomery, 2021). All control variables were measured in the baseline survey.

\hypertarget{analytical-strategy}{%
\subsubsection{Analytical Strategy}\label{analytical-strategy}}

@@

\hypertarget{limitations}{%
\subsubsection{Limitations}\label{limitations}}

@@

\hypertarget{appendix-appendix}{%
\appendix}
